An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-27 paper/ms.tex
update ms

% \begin{abstract}
%   Kepler\,1627A is a G8V star previously found to host a
%   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
%   selected for observation by the Kepler satellite because it is
%   nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
%   data from Gaia with rotation periods from TESS, we show that
%   Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
%   cluster.  Kepler\,1627Ab is therefore the youngest planet with a
%   precise age found to date by the main Kepler mission.
%   Analyzing the photometry, we find evidence that starspot crossings
%   distort the average transit shape, and induce transit timing
%   variations that suggest the planet is on a prograde orbit.  The 
%   flare arrival time distribution is most likely Poissonian, though
%   there may be hints ($<3$-$\sigma$) of additional flares at the
%   orbital and synodic periods.  In the future, spectroscopic studies
%   of \pn\ may help clarify the evolution of this mini-Neptune's orbit
%   and atmosphere.  More broadly, the combination of Gaia and TESS data
%   offers an opportunity to expand the census of age-dated planets; we
%   provide a literature concatenation of young stars to support this
%   effort.
% \end{abstract}

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the $35\pm XX$\,Myr old $\delta$~Lyr cluster.
  Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The precision of the Kepler
  photometry lets us detect two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are therefore well-poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of
molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster \citep[Stephenson-1;][]{stephenson_possible_1959,eggen_photometric_1968}.
The overlap between the Kepler field and the cluster is shown in
Figure~\ref{fig:skychart}.

A previous clustering analysis of the Gaia data
reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member of the
\cn\ \citep{kounkel_untangling_2019}.
Given the statistical validation of the close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018}, we therefore began by examining the age of
the star and its putative host cluster (Section~\ref{sec:cluster}). 
We find that the $\delta$ Lyr cluster is $\approx35$\,Myr old, and
that \sn\ is indeed a member of the cluster.
Focusing on the planet itself (Section~\ref{sec:planet}), we confirm
that despite a previously unreported binary companion, it is indeed a
mini-Neptune orbiting the roughly Sun-like primary star.
The presence of a correlation between transit timing variations and
the local light curve slope renders any other explanation implausible.
The broader implications for our ability to age-date a statistical
sample of planets are highlighted in closing (Section~\ref{sec:conc}).

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
    points are kinematically selected members of the $\delta$\,Lyr
    cluster.  Both Kepler (gray squares) and TESS (colored points)
    observed portions of the cluster.  Naked-eye stars ($m_{\rm
    V}<6.5$) are shown in gray; four of them (orange points) have
    their names annotated.  Kepler\,1627 (green star) was observed
    during the entirety of the Kepler mission, and has been observed
    by TESS for two lunar months to date.
    {\bf todo: move colorbar...}
		\label{fig:skychart}
	}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and Kepler
    rotations periods and dereddened Gaia colors, with the Pleiades
    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
    same comparison. {\bf todo: caption improve. fix figures.}.
   \label{fig:age}
	}
\end{figure*}

{\bf Jason Curtis todo}
We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

@@

\section{Kinematic Selection of $\delta$\,Lyr Cluster Members}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Galactic position and tangential velocities of the
			$\delta$\,Lyr cluster (also known as Theia 73 and Stephenson\,1).}  Points are
		candidate cluster members with $\varpi/\sigma_\varpi > 20$, reported
		to be in the group by \citet{kounkel_untangling_2019}.  We focus
		on stars in a small region (black points) in the kinematic
		vicinity of Kepler\,1627 (yellow star).  The other candidate cluster
		members (gray points) may or may not share the ages of the selected
		kinematic group.  The location of the Sun is ($\odot$) is shown.
		\label{fig:XYZvtang}
	}
\end{figure*}

Figure~\ref{fig:XYZvtang} shows members of the \cn\
reported
by \citet{kounkel_untangling_2019} to be in the group.
Galactic positions are determined and plotted only for stars with parallax
signal-to-noise exceeding 20.
The location of the Sun is shown on the plots.
The non-uniform ``clumps'' might be
an artifact of the data processing steps performed by \citet{kounkel_untangling_2019}.
We therefore only consider stars in the immediate kinematic group
around \sn. 
The tangential velocities relative to \sn\ are shown in the bottom right panel.
These are computed by assuming that every star has the same three-dimensional
spatial velocity as \sn, where we assume a systemic radial velocity of
$-16.7 \pm 0.2$\,\kms\ based on the reconnaissance spectra obtained by
A.~Howard on HIRES and D.~Latham on TRES.
The relevant projection effects are then taken into account,
as discussed by {\it e.g.}, \citet{Meingast2021} and {\bf L.~Bouma et al (2021, submitted)}.

>>>>

% 195 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
  with rotation periods from TESS, we show that Kepler\,1627 is a
  member of the {\bf $35\pm XX$\,Myr old} $\delta$~Lyr cluster.
  Kepler\,1627Ab is therefore the youngest planet with a precise age
  yet found by the main Kepler mission.  The Kepler
  photometry seems to show two anomalies.  First, unresolved
  starspot-crossing events distort the average transit shape, and
  induce transit timing variations that imply the planet is likely on
  a prograde orbit.  Second, although the flare arrival time
  distribution is probably Poissonian, there may be an excess of
  flares separated by the orbital and synodic periods, and linear
  combinations thereof.  The $\delta$~Lyr cluster is one of many
  stellar groups whose properties have been clarified using the Gaia
  data; many of the other new clusters also overlap with confirmed and
  candidate exoplanets.  Further studies that combine Gaia kinematics
  and TESS rotation periods are well-poised to expand the census of
  age-dated planets, and our understanding of planetary evolution.
\end{abstract}

@@

%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of
molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets, measuring their masses, and detecting their
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster
(Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
Figure~\ref{fig:skychart} shows the overlap between the Kepler field
and the cluster.

A previous HDBScan-based clustering analysis of the Gaia data found
that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member
of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
validation of close-in mini-Neptune
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we opted to examine the age of the star and its putative
host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
Lyr cluster is {\bf $\approx35$\,Myr old}, and that \sn\ is indeed a member
of the cluster.  Focusing on the planet itself
(Section~\ref{sec:planet}), we confirm that despite a previously
unreported binary companion, it is indeed a mini-Neptune orbiting the
roughly Solar-type primary star.  The presence of a correlation
between transit timing variations and the local light curve slope
renders any other explanation implausible.  We conclude by
highlighting some broader implications for our ability to age-date a
statistical sample of planets (Section~\ref{sec:conc}).

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
    points are kinematically selected members of the $\delta$\,Lyr
    cluster.  Both Kepler (gray squares) and TESS (colored points)
    observed portions of the cluster.  Naked-eye stars ($m_{\rm
    V}<6.5$) are shown in gray; four of them (orange points) have
    their names annotated.  Kepler\,1627 (green star) was observed
    during the entirety of the Kepler mission, and has been observed
    by TESS for two lunar months to date.
		\label{fig:skychart}
	}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f15a.pdf}
			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Left:} HR diagram of 
    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
    The \cn\ and IC\,2602 are approximately the same isochronal age.
    {\it Right:} TESS and Kepler rotations periods and dereddened Gaia
    colors, with the Pleiades
    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\bf todo: caption improve. fix
    figures.}.
   \label{fig:age}
	}
\end{figure*}

% paragraph 1: HR 
To assess the age of the cluster as well as our star of interest,
we used a combination of the isochronal, gyrochronal, and
lithium-based age-dating.

\subsection{HR Diagram}
We meaured the isochrone age using an empirical approach, and also
compared against theoretical models.
The left panel of Figure~\ref{fig:age} shows the HR diagram of 
stars in the \cn, the Pleiades, IC\,2602, and the field.
The stars shown in the \cn\ are selected based on their positions and
kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
Those from the Pleiades and IC\,2602 are adopted from
\citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.
% FIXME FIXME
We corrected for reddening following {\bf procedure X,Y,Z}.

The immediate impression from the left panel of Figure~\ref{fig:age}
is that the \cn\ and IC\,2602 share approximately the same isochronal
age, since they overlap.  In our exploration, we also compared against
$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and subsets
of the Sco OB2 association ($\approx10$-$15$\,Myr;
\citealt{damiani_stellar_2019}), and confirmed that the
pre-main-sequence M dwarfs were intermediate between these two
clusters.

% FIXME FIXME
We quantified this impression following the empirical fitting procedure
discussed by \citet{gagne_mutau_2020}.

\subsection{Rotation Periods}
{\bf Jason Curtis todo}
We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

@@

\section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
\label{app:kinematicselection}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Galactic position and tangential velocities of the
    $\delta$\,Lyr cluster.}
    Points are candidate cluster members with $\varpi/\sigma_\varpi >
    20$, reported to be in the group by
    \citet{kounkel_untangling_2019}.  We focus on stars in a small
    region (black points) in the kinematic vicinity of Kepler\,1627
    (yellow star).  The other candidate cluster members (gray points)
    may or may not share the ages of the selected kinematic group.
    The location of the Sun is ($\odot$) is shown.
		\label{fig:XYZvtang}
	}
\end{figure*}

Figure~\ref{fig:XYZvtang} shows stars reported by
\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
was cross-matched by \citealt{kounkel_untangling_2019} as being
``Stephenson 1''.  For our Figure, galactic positions were calculated
and plotted only for stars with parallax signal-to-noise exceeding 20.
The location of the Sun is shown on the plots.  The smattering of
reported cluster members (gray and black points) has a significantly
different structure that the cluster initially identified by
\citet{stephenson_possible_1959} and corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
might be part of a real structure, they could also be an artifact of
the data processing steps performed by
\citet{kounkel_untangling_2019}.  We therefore opted to only consider
stars in the immediate kinematic and spatial vicinity of \sn.  The
tangential velocities relative to \sn\ are shown in the bottom right
panel.  These are computed by assuming that every star has the same
three-dimensional spatial velocity as \sn, where we assume a systemic
radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the reconnaissance
spectra obtained by A.~Howard on HIRES and D.~Latham on TRES.  The
relevant projection effects are then taken into account, as discussed
by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
We performed the actual selection by then manually drawing lassos with
the interactive \texttt{glue} visualization tool
\citep{beaumont_2014_13866} in the four projections shown in
Figure~\ref{fig:XYZvtang}.  While bonafide members likely exist
outside of our selection region (and our selection also includes some
field star interlopers), our aim is to verify the existence of the
cluster in the vicinity of Kepler\,1627, and to measure its age.  The
procedure we have adopted enables both tasks.
